Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-16 paper_enahpc14/Sec3:Measurement-framework.tex
updates "Power-performance measurement framework" & "Evaluation" sections; update bibliography

\subsubsection{E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center of  ETH Zurich (CSCS) are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
(\url{http://riedonetworks.com/}),  that  enable  to monitor  and  log
power  consumption of  the  IT infrastructure  as  well as  constantly
analyze  line  voltage,  current,  power-factor,  frequency  with  1\%
accuracy.   Using reliable  narrowband  powerline communication  (PLC)
technology,  all metering  and power  quality data  from each  IPS are
centrally collected by the E3METER Data Concentrator, via the existing
power cables thus  avoiding the need for extra  cabling.  This data is
made  available  via SNMP,  HTTP,  TELNET  through  the built-in  Fast
Ethernet  port.   Time  synchronization  is guaranteed  by  using  NTP
servers.   Measured data  is accessed  through the  open  source Cacti
software including  the E3METER  Cacti Plugin to  scan the  entire PLC
network and monitor the power usage of individual rack in real-time.

\subsubsection{University of Hamburg}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with VampirTrace  and
Vampir,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  VampirTrace
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  The  server nodes  are connected  to power
measurement  devices that  account for  the  dissipated power/consumed
energy and  send the power data  to the tracing  server.  The attached
VampirTrace \pmlib plugin employs the client API that sends start/stop
primitives in order to gather captured data by the wattmeters onto the
tracing server,  where an  instance of the  \pmlib server  is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
the  power   data  from  the  tracing   server.   The  instrumentation
post-processing generates  the performance trace files  and the \pmlib
plugin inserts the power data into them.

In  addition  to the  power  measurements,  we  also account  for  the
resource utilization values  of the nodes: CPU load,  memory usage and
storage device utilization. % and network utilization.

We  run special  \pmlib  server  instances on  the  server nodes  that
retrieve these  values from the \texttt{proc}  file system (leveraging
the  \texttt{psutil} Python library).   Thus, \pmlib  plugin instances
running  with the  instrumented  application connect  with the  \pmlib
servers.   Finally,   using  the   Vampir   visualization  tool,   the
power-performance traces  can be easily  analyzed through a  series of
plots and statistics.

>>>>

\subsubsection{E3METER metering products}
Supercomputer  clusters considered  for our  experiments at  the Swiss
National Supercomputing Center of  ETH Zurich (CSCS) are equipped with
E3METER Intelligent  Power Strips (IPS)  and Monitors (IPM)  which are
high   quality   electricity  meters   released   by  Riedo   Networks
(\url{http://riedonetworks.com/}),  that  enable  to monitor  and  log
power  consumption of  the  IT infrastructure  as  well as  constantly
analyze  line  voltage,  current,  power-factor,  frequency  with  1\%
accuracy.   Using reliable  narrowband  powerline communication  (PLC)
technology,  all metering  and power  quality data  from each  IPS are
centrally collected by the E3METER Data Concentrator, via the existing
power cables thus  avoiding the need for extra  cabling.  This data is
made  available  via SNMP,  HTTP,  TELNET  through  the built-in  Fast
Ethernet  port.   Time  synchronization  is guaranteed  by  using  NTP
servers.   Measured data  is accessed  through the  open  source Cacti
software including  the E3METER  Cacti Plugin to  scan the  entire PLC
network and monitor  in real-time the power usage  of individual rack,
recorded in 5 minutes interval periods.

\subsubsection{University of Hamburg}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with Extrae  and
Paraver,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  Extrae
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  These  nodes are connected  to power 
measurement  devices that  account for  the dissipated power/consumed 
energy and  send the power data  to the tracing  server. The client, 
meanwhile, sends start/stop primitives in order to gather captured data
by the wattmeters onto the tracing server, where an  instance of the 
\pmlib server  is running.

Once COSMO-ART run is finished, a file containing the power profile is 
created using the  power   data  received from  the  tracing   server and
the  instrumentation post-processing generates  the performance trace files.
All these files are combined then in Paraver which allows us to visualize 
the performance trace and the power profile of COSMO-ART all together.

%Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
%the  power   data  from  the  tracing   server.   The  instrumentation
%post-processing generates  the performance trace files  and the \pmlib
%plugin inserts the power data into them.

%In  addition  to the  power  measurements,  we  also account  for  the
%resource utilization values  of the nodes: CPU load,  memory usage and
%storage device utilization. % and network utilization.

%We  run special  \pmlib  server  instances on  the  server nodes  that
%retrieve these  values from the \texttt{proc}  file system (leveraging
%the  \texttt{psutil} Python library).   Thus, \pmlib  plugin instances
%running  with the  instrumented  application connect  with the  \pmlib
%servers.   Finally,   using  the   Vampir   visualization  tool,   the
%power-performance traces  can be easily  analyzed through a  series of
%plots and statistics.

### 2014-04-16 paper_enahpc14/Sec4:Evaluation.tex
updates "Power-performance measurement framework" & "Evaluation" sections; update bibliography

\subsubsection{Sofware environment}
INT2LM  and the  COSMO-ART model  are  implemented in  Fortran 90  for
distributed  memory  parallel  computers  using  the  Message  Passing
Interface  (MPI) and is  purely MPI-parallel.   The software  stack on
both many-core  platforms was  controlled using the  modules framework
which gives  an easy and  flexible mechanism to  access to all  of the
CSCS  provided compilers,  tools  and applications.   For our  initial
benchmarking,  we opted  for  the GNU  compiler  (gcc/4.8.1 on  Monch,
gcc/4.8.2 on Pilatus) with the -O3 compiler flag as it generally gives
a  good  level  of optimization  and  the  code  runs faster  in  this
configuration  than when  compiled with  the intel  compiler (14.0.1).
Besides, we installed the  MPICH2 implementation of MPI (mvapich2/1.9)
as  well  as  the  commonly  used HDF5  (1.8.12)  and  NetCDF  (4.3.1)
libraries in favor of the  traditional GRIB library for the management
of extremely  large and complex data collections.   All computes nodes
have    an   operating   system    based   on    GNU/Linux   featuring
``2.6.32-358.11.1.el6.x86\_64''  and  ``3.0.101-0.15-default'' kernels
for Monch and Pilatus respectively.

>>>>

\subsubsection{Tintorrum (UJI)}
Tintorrum, which is installed at the Jaume I University (UJI),  is an 
heterogeneous cluster composed of 28 compute nodes. For our experiments
only a subset of these nodes were considered. This set is composed of 
16 nodes, each of which includes two Intel Xeon E5645 hexa-core processors at 
2.40 GHz connected via Infiniband QDR (Mellanox MTS3600 switch). 
Each core has 32 KB instruction and 32 KB data L1 caches and 
256 KB L2 cache. The 6 cores share a 12 MB L3 cache and the platform has 
288 GB of DDR3 1333 MHz.

\subsubsection{Sofware environment}
INT2LM  and the  COSMO-ART model  are  implemented in  Fortran 90  for
distributed  memory  parallel  computers  using  the  Message  Passing
Interface (MPI)  and are purely  MPI-parallel.  The software  stack on
both many-core  platforms was  controlled using the  modules framework
which gives  an easy and  flexible mechanism to  access to all  of the
CSCS  provided compilers,  tools  and applications.   For our  initial
benchmarking,  we opted  for  the GNU  compiler  (gcc/4.8.1 on  Monch,
gcc/4.8.2 on Pilatus) with the -O3 compiler flag as it generally gives
a  good  level  of optimization  and  the  code  runs faster  in  this
configuration  than when  compiled with  the intel  compiler (14.0.1).
Besides, we installed the  MPICH2 implementation of MPI (mvapich2/1.9)
as  well  as  the  commonly  used HDF5  (1.8.12)  and  NetCDF  (4.3.1)
libraries in favor of the  traditional GRIB library for the management
of extremely  large and complex data collections.   All computes nodes
have    an   operating   system    based   on    GNU/Linux   featuring
``2.6.32-358.11.1.el6.x86\_64''      kernel      in     Monch      and
``3.0.101-0.15-default'' kernel in Pilatus.
